---
title: "y04.1_Volcano"
author: "Trine Bertram Rasmussen"
date: '2022-06-22'
output: html_document
---

This script makes volcano plots and as MA-plot. I compare the data with unshrunken LFC to 
data with shrunken LFC with these plots. It does lower the LFC a lot but doesnt seem to
affect the significant DEGs which is good. 
Lets continue the analysis with shrunken LFC.

```{r setup, include=FALSE}

library("ggplot2") #for plotting
library("DESeq2") # for DESeq analysis
library("EnhancedVolcano")
library(ggpubr)
library("patchwork") # for merging the three volcano plots

setwd("./")

#made in 03_DESeqObject
dds <- readRDS("R_outputs/dds.rds") #DESeq object
res <- readRDS("R_outputs/res.rds") #result table from DESeq object

# contrasts
Fa_C_unshrunken <- readRDS("R_outputs/Fa_C_unshrunken.rds")
Aa_C_unshrunken <- readRDS("R_outputs/Aa_C_unshrunken.rds")
Aa_Fa_C_unshrunken <- readRDS("R_outputs/Aa_Fa_C_unshrunken.rds")

Fa_C <- readRDS("R_outputs/Fa_C.rds")
Aa_C <- readRDS("R_outputs/Aa_C.rds")
Aa_Fa_C <- readRDS("R_outputs/Aa_Fa_C.rds")

```


```{r Volcano plots unshrunken, fig.height=10, fig.width=12}
theme_set(theme_pubr())

pCutoff = 0.05
FCcutoff = 1.0

#Interesting genes
chosen_genes <- c("YOR388C","YNR034W-A","YOL052C-A" , "YPL276W", "YLR258W","YMR169C","YPL275W")

VolcanoPlot1 <- EnhancedVolcano(Fa_C_unshrunken,
                                        lab = rownames(Fa_C_unshrunken),
                                        selectLab = chosen_genes,
                                        #selectLab = NA,
                                        x = 'log2FoldChange',
                                        y = 'padj',
                                        title = "Furfural",
                                        subtitle = bquote(italic(" ")),
                                        legendPosition = 'top',
                                        pCutoff = pCutoff,
                                        FCcutoff = FCcutoff,
                                        drawConnectors = TRUE,
                                        widthConnectors = 1.0,
                                        ylim = c(0,350),
                                        xlim = c(-5,8)
                                        )
                                        

VolcanoPlot2 <- EnhancedVolcano(Aa_C_unshrunken,
                                        lab = rownames(Aa_C_unshrunken),
                                        selectLab = chosen_genes,
                                        #selectLab = NA,
                                        x = 'log2FoldChange',
                                        y = 'padj',
                                        title = "Acetic acid",
                                        subtitle = bquote(italic(" ")),
                                        legendPosition = 'top',
                                        pCutoff = pCutoff,
                                        FCcutoff = FCcutoff,
                                        drawConnectors = TRUE,
                                        widthConnectors = 1.0,
                                        ylim = c(0,350),
                                        xlim = c(-5,8)
                                        )

VolcanoPlot3 <- EnhancedVolcano(Aa_Fa_C_unshrunken,
                                        lab = rownames(Aa_Fa_C_unshrunken),
                                        selectLab = chosen_genes,
                                        #selectLab = NA,
                                        x = 'log2FoldChange',
                                        y = 'padj',
                                        title = "Acitid acid and furfural",
                                        subtitle = bquote(italic(" ")),
                                        legendPosition = 'top',
                                        pCutoff = pCutoff,
                                        FCcutoff = FCcutoff,
                                        drawConnectors = TRUE,
                                        widthConnectors = 1.0,
                                        ylim = c(0,350),
                                        xlim = c(-5,8)
                                        )


VolcanoPlot1 + VolcanoPlot2 + VolcanoPlot3
Combined <- VolcanoPlot1 + VolcanoPlot2 + VolcanoPlot3

ggsave(
  "Volcano_Combined.pdf",
  plot = Combined,
  path = "R_outputs/",
  height = 7,
  width = 12,
  dpi = 1000
)
```

```{r Volcano plots shrunk, fig.height=10, fig.width=12}
theme_set(theme_pubr())

pCutoff = 0.05
FCcutoff = 1.0

#Interesting genes
chosen_genes <- c("YOR388C","YNR034W-A","YOL052C-A" , "YPL276W", "YLR258W","YMR169C","YPL275W")

VolcanoPlot1 <- EnhancedVolcano(Fa_C,
                                        lab = rownames(Fa_C),
                                        selectLab = chosen_genes,
                                        #selectLab = NA,
                                        x = 'log2FoldChange',
                                        y = 'padj',
                                        title = "Furfural",
                                        subtitle = bquote(italic(" ")),
                                        legendPosition = 'top',
                                        pCutoff = pCutoff,
                                        FCcutoff = FCcutoff,
                                        drawConnectors = TRUE,
                                        widthConnectors = 1.0,
                                        ylim = c(0,350),
                                        xlim = c(-5,8)
                                        )
                                        

VolcanoPlot2 <- EnhancedVolcano(Aa_C,
                                        lab = rownames(Aa_C),
                                        selectLab = chosen_genes,
                                        #selectLab = NA,
                                        x = 'log2FoldChange',
                                        y = 'padj',
                                        title = "Acetic acid",
                                        subtitle = bquote(italic(" ")),
                                        legendPosition = 'top',
                                        pCutoff = pCutoff,
                                        FCcutoff = FCcutoff,
                                        drawConnectors = TRUE,
                                        widthConnectors = 1.0,
                                        ylim = c(0,350),
                                        xlim = c(-5,8)
                                        )

VolcanoPlot3 <- EnhancedVolcano(Aa_Fa_C,
                                        lab = rownames(Aa_Fa_C),
                                        selectLab = chosen_genes,
                                        #selectLab = NA,
                                        x = 'log2FoldChange',
                                        y = 'padj',
                                        title = "Acitid acid and furfural",
                                        subtitle = bquote(italic(" ")),
                                        legendPosition = 'top',
                                        pCutoff = pCutoff,
                                        FCcutoff = FCcutoff,
                                        drawConnectors = TRUE,
                                        widthConnectors = 1.0,
                                        ylim = c(0,350),
                                        xlim = c(-5,8)
                                        )


VolcanoPlot1 + VolcanoPlot2 + VolcanoPlot3
Combined <- VolcanoPlot1 + VolcanoPlot2 + VolcanoPlot3

ggsave(
  "Volcano_Combined_shrunkLFC.pdf",
  plot = Combined,
  path = "R_outputs/",
  height = 7,
  width = 12,
  dpi = 1000
)

```
CONCLUSION: The lfcShrink() function did not change what genes that we recognised before as important. So that is very nice! From now on, we should use the results with shrunken values.


```{r plotCounts}
par(mfrow=c(2,3))
plotCounts(dds, gene="YOR388C", intgroup="condition")
plotCounts(dds, gene="YPL276W", intgroup="condition")
plotCounts(dds, gene="YPL275W", intgroup="condition")

plotCounts(dds, gene="YMR169C", intgroup="condition")
plotCounts(dds, gene="YMR320W", intgroup="condition")
plotCounts(dds, gene="YPL277C", intgroup="condition")

```



## show the effect of the shrinked lfc for our data

We can see that the LFC for the lowly expressed genes is corrected with the lcfShrink().
However, it does not seem to affect any significant genes, as indicated in the volcano plot.
```{r MAplot, fig.height=7, fig.width=7}
par(mfrow=c(2,1))
plotMA(Fa_C_unshrunken,
       colNonSig = "black")
plotMA(Fa_C,
       colNonSig = "black")
```


