---
title: "PCR and adonis"
author: "Trine Bertram Rasmussen"
date: '2022-06-23'
output: html_document
---
Herein, we explore the groupings on the samples based on their gene expression profile. 
The PCA plot shows that the gene expression is very different among the different conditions. 
The ADONIS test is a statistical test that measures the significance of this
grouping observed in the PCA plot, which is very significant. 


```{r}
library("ggplot2") #for plotting
library("DESeq2") # for DESeq analysis
library("vegan") #to make adonis analysis

source("y00_functions.R")
# made in 03_DESeqObject
dds <- readRDS("R_outputs/dds.rds") #DESeq object
res <- readRDS("R_outputs/res.rds") #result table from DESeq object

metaData <- read_clean_metaData() # metadata about samples
countData <- read_clean_countData() # countdata about samples
```


#Explore raw data with PCA plots and test if the groupings are significant with adonis

```{r}


# First we need to transform the raw count data
# vst function will perform variance stabilizing transformation

vsdata <- vst(dds, blind=FALSE)


PCA_top500 <- plotPCA(vsdata, intgroup="condition")

source("y00_functions.R") #to get theme_Publication()
PCA_top500_save <- PCA_top500 + theme_Publication() + 
  theme(strip.text.x = element_blank(), 
        strip.text.y = element_blank(), 
        axis.title.x = element_text(size = 16), axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 16, hjust=1), axis.text.y = element_text(size = 16, hjust = 1),
        legend.key.size = unit(0.8, "cm"), 
        legend.text=element_text(size=15),
        legend.position = c(0.80, 0.85),legend.background = element_rect(fill = "white", color = "black")) + 
  scale_colour_Publication(name="Experimental\nCondition",
                           breaks=c("C", "Fa", "Aa", "Aa_Fa"),
                           labels=c("Control", "Furfural", "Acidic acid", "Acidic acid + furfural"))+     
  geom_point(size = 4) 

#geom_text(mapping = aes(label = name), size = 4.5, vjust=1.5)

PCA_top500_save 


ggsave(
  "PCA_top500_save.pdf",
  plot = PCA_top500_save,
  path = "R_outputs/",
  height = 7,
  width = 12,
  dpi = 1000
)

```


#ADONIS test

```{r}
# Make an ANOVA test/ ADONIS test
# Tutorial: https://www.rdocumentation.org/packages/vegan/versions/2.4-2/topics/adonis

# format countData for adonis
countData_adonis <-t(countData[,-1])
colnames(countData_adonis) <- countData[, 1]

ncol(countData_adonis) #11126
set.seed(123456789)
adonis_condition <- adonis2(countData_adonis ~ condition, data = metaData, permutations = 10000) 
# p =9.999e-05 ***
# The p value is HIGHLY significant.
```


